---
title: "Analysis of Inflation and Wages"
output: html_document
---

Now that you have established that Inflation increasing at a higher rate than wages. Lets see how bad of an issue this seems to be.

## Importing Data

```{r Import, message=FALSE}
library(readxl)
DatasetForRProject <- read_excel("DatasetForRProject.xlsx")
```

## Descriptive Statistics

```{r DescriptiveStat, message=FALSE}
summary(DatasetForRProject)
```

The General Average Wage has much higher measures of spread than median measures of spread.
This may indicate that there are large outliers influencing the average wage to be higher.
If we look at wages-by-degree, it seems like having a degree is much more helpful for wages than
not having a degree.

```{r LineGraphs, message=FALSE}
library(ggplot2)
ggplot(data=DatasetForRProject, aes(x=Year, y=GeneralMedianWage, group=1)) +
  geom_line()+
  geom_point()
ggplot(data=DatasetForRProject, aes(x=Year, y=GeneralAverageWage, group=1)) +
  geom_line()+
  geom_point()
```

This seems like average has less point-to-point variation than the median wage category. Overall, it seems like Wages increase over time and sharply over the past 20 years.

## Correlations

```{r InitialCorr, message=FALSE}
library(Hmisc)
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
res2<-rcorr(as.matrix(DatasetForRProject[,1:9]))
Corr <-flattenCorrMatrix(res2$r, res2$P)
Corr[ , c("row", "column","cor")]    
```

There are high correlations between Consumer Price Index and median wage, average wage, Bachelor's 
Degree, and Advanced Degree. There was not a high correlation with wages for some college, High School,
and less than High School. This is highly concerning. This means that over time, individuals who
receive less than a Bachelor's degree are not keeping up with inflation as much as a college degree. 
Also, the correlation is negative for people who do not have a high school diploma. This
indicates that as the Consumer Price Index increases over time, people with a high school diploma
actually have their wages decrease. 

## Conclusions

The results from the correlations were extremely interesting and shed light on our situation in the
United States. Individuals who have a College Degree seem to have wages that keep up with inflation.
Those who do not have a degree, but have a high school education, increase over time. But not at the 
same pace as inflation. The worth of their wages are slowly eroding over time. For those without a 
high school diploma, their wages actually decrease as inflation increases.

With inflation in the United States at a high in recent history, it seems that those with a college
diploma will be better off than those without.